---
title: "Texas Housing"
author: "Intro to Data Science"
date: "Week 04"
output: html_notebook
---

- An opportunity to practice some grouping, sorting, and mutating.

# Task 00 - Setup

- Complete the code chunk below to start tonight's lab.
- Then, run the code-chunk below to get started.
- Look at tonight's lecture if you feel stumped.
- This task is not connected to any of the questions on Canvas, although obviously completion is a prerequisite.

```{r}
library(knitr)
library(rio)
library(tidyverse)

## Disables scientific notation.
options(scipen=999)

## How do we import our data?
txhousing <- read_csv("data/txhousing.csv")
```

# Data

- Look at our data. What is the grain?

```{r}
txhousing
```



# Abilene

## Task 01

- **Question:** How many rows of data do we have about Abilene Texas?
- **Answer:** 

Helpful Hints:

- Think about `filter()`
- Remember to wrap your string in quotes.
- You must use `==` to perform a Boolean test!

```{r}
txhousing |> filter(city == "Abilene")

txhousing |> count(city)
```

## Task 02

- **Question:** In what year/month did Abilene have the highest volume of sales?
- **Answer:**

Helpful Hints:

- This is easier if you used `filter()` for Task 01.
- If you did, you only need to add one thing to your Task 01 code to answer this question.

```{r}
txhousing |> filter(city == "Abilene") |> arrange(desc(volume))
```

## Task 03

- **Question:** How many houses (total) were sold in Abilene in 2010?
- **Answer:**

Helpful Hints:

- To solve this, you will have to change the grain of the data.
- You can solve this problem with either `filter` or `group_by`, although `filter` is arguably more elegant.
    - To have more than one filter test, separate them with a comma.
    - This will treat them like "and" statements.
    - Alternatively, you could use the `&` symbol.
- You cannot avoid `summarize`.
- The function `sum` may be useful.
- Type `?sum` into the Console if you would like to review the documentation on this function.

```{r}
txhousing |>
  filter(city == "Abilene", year == 2010) |>
  summarize(
    total_sales = sum(sales),
    avg_sales = mean(sales),
    total_listings = sum(listings)
    )
```

## Task 04

- **Question:** Which city had more total homes sold after 2013, Abilene or Austin?
- **Answer:**

```{r}
txhousing |>
  filter(year > 2013) |>
  group_by(city) |>
  summarize(
    total_sales = sum(sales, na.rm = TRUE)
    )
```

# I'm leaving Abilene, tonight

This is for all the Carolyn Carter fans out there (can't find the original).

<iframe width="560" height="315" src="https://www.youtube.com/embed/C2oUbDd3Jn8" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Task 05

- **Question:** How many homes were sold in each city in each year?
- **Answer:**

Helpful Hints:

- Create a new data frame with a city/year grain.
- Save this as a new data frame called txcities.
- Make sure you look at your data to confirm that you have transformed it correctly.
- You can copy/paste the code from your code-chunk to Canvas.

```{r}
txcities <- 
  txhousing |>
  group_by(city, year) |>
  summarize(
    total_sales = sum(sales, na.rm = TRUE)
    )
txcities
```

## Task 06

- **Question:** Plot the data from txcities, but only plot those cities with more than 10,000 homes sold in a year. Otherwise, your plot will be cluttered.
- **Answer:** Upload to Canvas.

Helpful Hints:

- Yes, you will need to filter txcities before plotting it.
    - You can handle this in a single statement or you can create a third data set.
- Since this is a time series and not a scatter plot, you will need to draw a line, not points.
    - `geom_line()` is your friend here.
    - Otherwise, the x/y axes should be familiar.
    - To draw a separate line for each city, assign the city column to color in your `aes()`.
- Download your plot from here, to upload to Canvas.

```{r}
txcities |>
  filter(total_sales > 10000) |>
  ggplot(aes(x = year, y = total_sales, color = city)) +
  geom_line()
```

## Task 07

- **Question:** I hope you got a graph that shows some high points in 2006, followed by a steep decline. Can you explain what happened between 2008 and 2010 in the housing market? (Hint: this trend was much bigger than the Texas housing market.)
- **Answer:** The 2008 economic crash.

## Task 08

- **Question:** Plot the median house value of homes sold in Houston from 2006 to 2012.
- **Answer:** Upload to Canvas.

Helpful Hints:

- Another time series, another `geom_line`.
- Your x axis must be the `date` column.
- You will need to filter, you will not need to transform this data set.
- Download the plot from RStudio and upload to Canvas.

```{r}
txhousing |>
  filter(city == "Houston", year >= 2006, year <= 2012) |>
  ggplot(aes(x = date, y = median, color = city)) +
  geom_line()
```

## Task 09

- **Question:** On average, which month has the highest number of sales in Houston?
- **Answer:** 

```{r}
txhousing |>
  filter(city == "Houston") |>
  group_by(month) |>
  summarize(avg_sales = mean(sales)) |>
  arrange(desc(avg_sales))

```

## Task 10

- **Question:** The summary statistic representing the value per home sold is the median value. Why not use the average?
- **Answer:** 

Helpful Hint:

- [Wikipedia: Median](https://en.wikipedia.org/wiki/Median)