---
title: "Texas Housing"
author: "Intro to Data Science"
date: "Week 04"
output: html_notebook
---

An opportunity to practice some grouping, sorting, and mutating.

# Setup

Run the code-chunk below to load everything you need for tonight's lab.

```{r}
rm(list = ls()) ## Deletes anything you may have created earlier, freeing up RAM.
library(knitr)
library(tidyverse)

## Show all the code from here down.
opts_chunk$set(echo = TRUE)

## Disables scientific notation.
options(scipen=999)

## This is a subset of the data from the gapminder presentation above.
data(txhousing)
txhousing
```


Each row of data information on a single month's worth of sales. So, we could say this is data is at the city/month grain.

- For more help on the data `?txhousing` in the console pane.

# Abilene

## Task 01

How many rows of data do we have about Abilene Texas?

- Think about `filter()`
- Remember to wrap your string in quotes.
- You must use `==` to perform a Boolean test!

```{r}
## YOUR CODE HERE!

```

## Task 02

In what year/month did Abilene have the highest volume of sales?

- This is easier if you used `filter()` for Task 01.
- If you did, you only need to add one thing to your Task 01 code to answer this question.

```{r}
## YOUR CODE HERE!

```

## Task 03

How many houses (total) were sold in Abilene in 2010?

- To solve this, you will have to change the grain of the data.
- You can solve this problem with either `filter` or `group_by`, although `filter` is arguably more elegant.
    - To have more than one filter test, separate them with a comma.
    - This will treat them like "and" statements.
    - Alternatively, you could use the `&` symbol.
- You cannot avoid `summarize`.
- The function `sum` may be useful.
- Type `?sum` into the Console if you would like to review the documentation on this function.

```{r}
## YOUR CODE HERE!

```

## Task 04

Which city had more total homes sold after 2013, Abilene or Austin?

```{r}
## YOUR CODE HERE!

```

# I'm leaving Abilene, tonight

This is for all the Carolyn Carter fans out there (can't find the original).

<iframe width="560" height="315" src="https://www.youtube.com/embed/C2oUbDd3Jn8" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Task 05

Calculate the total number of homes sold for all cities in each year.

- This will make a new data frame with a cit/year grain.
- Save this as a new data frame called txcities.
- Make sure you look at your data to confirm that you have transformed it correctly.
- You can copy/paste the code from your code-chunk to Canvas.

```{r}
## YOUR CODE HERE!

```

## Task 06

Plot the data from txcities, but only plot those cities with more than 10,000 homes sold in a year. Otherwise, your plot will be cluttered.

- Yes, you will need to filter txcities before plotting it.
    - You can handle this in a single statement or you can create a third data set.
- Since this is a time series and not a scatter plot, you will need to draw a line, not points.
    - `geom_line()` is your friend here.
    - Otherwise, the x/y axes should be familiar.
    - To draw a separate line for each city, assign the city column to color in your `aes()`.
- Download your plot from here, to upload to Canvas.

```{r}
## YOUR CODE HERE!

```

## Task 07

I hope you got a graph that shows some high points in 2006, followed by a steep decline. Can you explain what happened between 2008 and 2010 in the housing market? (Hint: this trend was much bigger than the Texas housing market.)

## Task 08

Plot the median house value of homes sold in Houston from 2006 to 2012.

- Another time series, another `geom_line`.
- Your x axis must be the `date` column.
- You will need to filter, you will not need to transform this data set.
- Download the plot from RStudio and upload to Canvas.

```{r}
## YOUR CODE HERE!

```

## Task 09

On average, which month has the highest number of sales in Houston?

```{r}
## YOUR CODE HERE!

```


## Task 10

The summary statistic representing the value per home sold is the median value. Why not use the average?

- [Wikipedia: Median](https://en.wikipedia.org/wiki/Median)